---
title: "Lab 05 - Wrangling spatial data"
author: "Heather Hawkins"
date: "02/08/23"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{R dnak}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are only 3 dennys in Alaska

### Exercise 2
```{R la-ak}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are only 2 La Quintas in Alaska

### Exercise 3

There are 6 pairing b/w all Denny's and La Quinta locations in Alasaka, so we will have to calculate 6 distances. 
```{R Distance}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

...

### Exercise 4

There are 6 observations in the new data frame.

There are 11 variables. address.x, city.x, state, zip.x, longitude.x,latitude.x, address.y, city.y, zip.y, address.y, city.y, zip.y, longitude.y, latitude.y

x= dn, y= la

...

### Exercise 5
What function from the tidyverse do we use the add a new variable to a data frame while keeping the existing variables?

We use function()
```{R haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) 
}
```

...

### Exercise 6

```{R Distances}
dn_lq_ak <- dn_lq_ak %>% mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))
```
...
### Exercise 7
Calculate the minimum distance between a Denny’s and La Quinta for each Denny’s location. To do so we group by Denny’s locations and calculate a new variable that stores the information for the minimum distance.


```{R min-distance}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```
...
### Exercise 8
Describe the distribution of the distances Denny’s and the nearest La Quinta locations in Alaska. Also include an approipriate visualization and relevant summary statistics.

```{R Final-Distance}
view(dn_lq_ak_mindist)

dn_lq_ak_mindist %>%
  ggplot (aes(x=closest))+
  geom_bar()+
  labs(title = "Closest Distance between a Denny's and La Quinta in Alaska",
       x="Distance in kilometers",
       y="Number of locations")
```

All 3 Denny's are 6 km (or under) away from a la quinta
...
### Exercise 9

```{R NC}

# Filtering for NC
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)

lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)

#Joining Data Frames

dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

#Calculating Distance

dn_lq_nc <- dn_lq_nc %>% mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))


# Calculating Minimum Distance

dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#Visualizing 

view(dn_lq_nc_mindist)

dn_lq_nc_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram()+
  labs(title = "Closest Distance between a Denny's and La Quinta in NC",
       x="Distance in kilometers",
       y="Number of locations")

  
summary(dn_lq_nc_mindist)

```
...
### Exercise 10
Repeat the same analysis for Texas.
```{R TX}

# Filtering for TX
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)

lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)

#Joining Data Frames

dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx

#Calculating Distance

dn_lq_tx <- dn_lq_tx %>% mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))


# Calculating Minimum Distance

dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#Visualizing 

view(dn_lq_tx_mindist)

dn_lq_tx_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram()+
  labs(title = "Closest Distance between a Denny's and La Quinta in Texas",
       x="Distance in kilometers",
       y="Number of locations")

  
summary(dn_lq_tx_mindist)

```
...
### Exercise 11
Repeat the same analysis for a state of your choosing, different than the ones we covered so far.
CA 
```{R CA}

# Filtering for CA
dn_ca <- dennys %>%
  filter(state == "CA")
nrow(dn_ca)

lq_ca <- laquinta %>%
  filter(state == "CA")
nrow(lq_ca)

#Joining Data Frames

dn_lq_ca <- full_join(dn_ca, lq_ca, by = "state")
dn_lq_ca

#Calculating Distance

dn_lq_ca <- dn_lq_ca %>% mutate(distance = haversine(longitude.x,latitude.x,longitude.y,latitude.y))


# Calculating Minimum Distance

dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#Visualizing 

view(dn_lq_ca_mindist)

dn_lq_ca_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram()+
  labs(title = "Closest Distance between a Denny's and La Quinta in California",
       x="Distance in kilometers",
       y="Number of locations")

  
summary(dn_lq_ca_mindist)

```
...

### Exercise 12

##AK:
Number of Denny's: 3, Number of La Quintas: 2, Minimum Distance: 2 km, Average Distance: 6.5 km

##NC:
Number of Denny's: 28, Number of La Quintas: 12, Minimum Distance: 1.8 km, Average Distance: 53.46 km

##TX:
Number of Denny's: 200, Number of La Quintas: 237, Minimum Distance:  0.02 km, Average Distance: 3.37 km

##CA:
Number of Denny's: 403, Number of La Quintas: 56, Minimum Distance:  0.02 km, Average Distance: 11.90 km


Mitch Hedberg's joke is MOST likely to hold true in Texas. This is because the average distance between a Denny's and La Quinta is 3.37 km. This is the lowest among all of the states. Additionally, TX has the most Denny's (200), the second close number of La Quintas (237), and a min distance of 0.02 km.

Although Alaska Average. Distance is 6.5 km (closer to TX than the other states), there are only 3 Denny's and 2 La Quintas. In comparison, CA has 11.9 km, but with 403 Denny's and 56 La Quinta's. 

...

